---
title: "Wind_speeds"
author: "Whitney"
date: "11/15/2021"
output:
  html_document: default
  pdf_document: default
---

##Install packages

```{r}
library(tidyr)
library(ggplot2)
library(dplyr)
```

##Read in data

```{r}
setwd("/Users/Owner/Desktop/R Practice/")
sampling_df <- read.csv("/Users/Owner/Desktop/R Practice/productivity_data_calculations.csv")
weather <- read.csv("/Users/Owner/Desktop/R Practice/2019_PASI.csv", skip=6)
weather <- weather[-c(1),]
```

##Combine date and time into a single column in data file

```{r}
colnames(sampling_df)
sampling2 <- sampling_df %>% 
  unite(date_time, Date, Init_time_hh.mm, sep = " ", remove = FALSE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

##Convert to POSIXct for Alaska time zone for sampling data frame

```{r}
sampling2$date_time_local <- as.POSIXct(sampling2$date_time, tz="US/Alaska",
                            format="%m/%d/%Y %H:%M:%S")
head(sampling2$date_time_local)
```

##Convert to POSIXct for weather file

```{r}
weather$date_time_UTC <- as.POSIXct(weather$Date_Time, tz="UTC",format="%m/%d/%Y %H:%M")
head(weather$date_time_UTC)
```

##Convert sampling file to UTC

```{r}
sampling2$date_time_UTC <- sampling2$date_time_local
attr(sampling2$date_time_UTC, "tzone") <- "UTC"
head(sampling2[,c("date_time_local","date_time_UTC")])
```

##Stick the wind speed from the weather data frame in a new column in the sampling data frame based on the closest time value

```{r}
sampling2$wind_speed = NA
for(i in 1:nrow(sampling2)) {
  rowindx = which.min(abs(sampling2$date_time_UTC[i] - weather$date_time_UTC))
  sampling2$wind_speed[i] = weather$wind_speed_set_1[rowindx] 
}
```

##Produce a histogram of wind speeds associated with the sampling times

```{r}
sampling2$wind_speed = as.numeric(sampling2$wind_speed)
hist(sampling2$wind_speed, main = 'Histogram of Sampling Wind Speeds (m/s)', 
     xlab = 'Wind speed', las=1)
```
  
##Calculate summary statistics for the wind speeds associated with the sampling times: mean, SD, quartiles. Produce a set of labeled numbers in the output document.

```{r}
mean(sampling2$wind_speed)
sd(sampling2$wind_speed)
quantile(sampling2$wind_speed, 0.25)
quantile(sampling2$wind_speed, 0.5)
quantile(sampling2$wind_speed, 0.75)
```

##Produce a line plot of the wind speeds for all of 2019, with date on the x-axis

```{r}
weather <- weather %>%
  rename(wind_speed = wind_speed_set_1)
weather$wind_speed = as.numeric(weather$wind_speed)
ggplot(weather, aes(x=date_time_UTC, y=wind_speed)) +
  geom_line()
```

##Calculate summary statistics for the 2019 wind speed data (weather). What were the mean, SD, and quartiles of wind speeds for the whole year? Produce a set of labeled numbers in the output document. 

```{r}
mean(weather$wind_speed, na.rm=TRUE)
sd(weather$wind_speed, na.rm=TRUE)
quantile(weather$wind_speed, 0.25, na.rm=TRUE)
quantile(weather$wind_speed, 0.5, na.rm=TRUE)
quantile(weather$wind_speed, 0.75, na.rm=TRUE)
```



